---
title: "Predict Outcomes from the amount of missingness"
author: Trang Le, Emily Getzen
output: html_document
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "htmls") })
params:
  data_dir: "/users/emily/Downloads"
  package_dir: "/users/emily/Documents/4ce/missing-data-4ce"
  results_file: "/penn_results-predict-te.Rdata"
  dateFormat: "%d-%b-%y"
---
```{r setup}
source('R/utils.R')
required_pkgs <- c('tidyverse', 'lubridate', 'stm', 'RColorBrewer','beepr','RcppRoll','purrr','mSTEM','Rcpp','MCMCpack')
install_required(required_pkgs)


library(tidyverse)
library(lubridate)
library(stm)
library(RColorBrewer)
library(beepr)
library(RcppRoll)
library(purrr)
library(mSTEM)
library(Rcpp)
library(MCMCpack)
```

```{r warning=FALSE, message=FALSE}

#replace my_dir with the directory that contains your files

my_dir <- params$data_dir

#load beta from NWU: located in package directory
load(paste(params$package_dir,"/public-data/beta_matrix2_nwu.Rdata",sep=""))
apply_betas = beta_mat2


theme_set(theme_bw() +
  theme(
    legend.title = element_blank(),
    panel.grid.minor = element_blank()
  ))
```

## Read in files 
and convert to wide format

```{R message=FALSE}

# 4CE long format Labs Data

patient_obs <- read_csv(file.path(my_dir, "LocalPatientObservations.csv"),
  col_types = list(patient_num = readr::col_character())
)

# Code, Descriptions and Ranges
lab_mapping <- read_csv("public-data/loinc-map.csv")
lab_bounds <- read_csv("public-data/lab_bounds.csv")
lab_names <- lab_bounds$short_name

#Read in local patient summary

demo_raw <-
  readr::read_csv(
    file.path(my_dir, "LocalPatientSummary.csv"),
    col_types = list(patient_num = readr::col_character()),
    na = "1900-01-01"
  ) %>%
  mutate(
    across(ends_with("_date") & where(is.character), lubridate::mdy),     
    last_discharge_date = if_else(
      !is.na(death_date) & death_date < last_discharge_date,
      death_date,
      last_discharge_date
    )
  )



```



```{r}

#Read in thrombotic icd codes. Add column "truncated code" for first three characters of each code. Extract unique truncated codes.

thrombo_codes <- read_csv("https://raw.githubusercontent.com/covidclinical/Phase2.1AKIRPackage/ac19716a4586f45c398728fcd821ca9d5baffe45/FourCePhase2.1AKI/data-raw/thromb_icd_code.csv") %>%
  mutate(truncated_code = substr(icd_code, 1, 3)) %>%
  pull(truncated_code) %>%
  unique()


#Generate list of neuro codes (obtained from neuro group)

neuro_codes <- c("R41","R27","R42","G44","G03","G04","G72","M60","G61","G65","R43","G93","F29","G40","G45","G46","I60","I61","I62","I67","H54","40","298","307","320","321","322","323","330","331","339","345","348","357","359","369","430","431","432")

ARDs_codes <- c("J80","518")

```
```{r}

#Isolate patients with thrombotic events and remove the days_since_admission column. 

thrombo_patients <- patient_obs %>%
  filter(
    concept_type == "DIAG-ICD10",
    concept_code %in% thrombo_codes,
    days_since_admission >= 0
  ) %>%
  dplyr::select(-days_since_admission) %>%
  distinct()


neuro_patients <- patient_obs %>%
  filter(
    concept_type == "DIAG-ICD10",
    concept_code %in% neuro_codes,
    days_since_admission >= 0
  ) %>%
  dplyr::select(-days_since_admission) %>%
  distinct()

ARDs_patients <- patient_obs %>%
  filter(
    concept_type == "DIAG-ICD10",
    concept_code %in% ARDs_codes,
    days_since_admission >= 0
  ) %>%
  dplyr::select(-days_since_admission) %>%
  distinct()

#Isolate information for unique patients. Sum individual patients with thrombotic events. 

demo_df <- demo_raw %>%
  mutate(
    TE = patient_num %in% unique(thrombo_patients$patient_num),
    deceased = deceased == 1,
    severe = severe == 1
  )
demo_df$TE %>% sum()


demo_df <- demo_df %>%
  mutate(
    Neuro = patient_num %in% unique(neuro_patients$patient_num),
    deceased = deceased == 1,
    severe = severe == 1
  )
demo_df$Neuro %>% sum()

demo_df <- demo_df %>%
  mutate(
    ARDs = patient_num %in% unique(ARDs_patients$patient_num),
    deceased = deceased == 1,
    severe = severe == 1
  )
demo_df$ARDs %>% sum()

#Generate a table that contains information about each lab for each individual patient. Each observation is identified by their patient number and days since admission, and each lab name and value is included. If there is more than one lab value for that identifier, they are averaged.

patient_obs_wide <- patient_obs %>%
  left_join(lab_bounds, by = c("concept_code" = "LOINC")) %>%
  dplyr::select(-concept_code) %>%
  filter(!is.na(short_name)) %>% 
  pivot_wider(
    id_cols = c(patient_num, days_since_admission),
    names_from = short_name,
    values_from = value,
    values_fn = mean
  )

```


## Unsupervised learning: LDA

```{r}
upper_day <- 9
err_xgbs <- vector("numeric")
err_svms <- vector("numeric")
all_evals <- list()

#filter out observations with days since admission >= a threshold (upper_day, in this case 9 days) and days since admission  <= 0. Take out the days_since_admission column. Obtain number of lab values for each TE patient.

missing_by_patient <- patient_obs_wide %>%
  filter(
    days_since_admission <= upper_day, # 10-day window
    days_since_admission >= 0
  ) %>%
  group_by(patient_num) %>%
  summarise(across(-days_since_admission, function(x) {
    sum(!is.na(x))
  })) %>%
  left_join(demo_df, by = "patient_num")

ct = 0
if ("Troponin_high" %in% colnames(missing_by_patient)){
  if ("Troponin_normal" %in% colnames(missing_by_patient)){
    ct = ct + 1
  }else{
  colnames(missing_by_patient)[which(colnames(missing_by_patient)=="Troponin_high")] <- "Troponin"
  }
}

if ("Troponin_normal" %in% colnames(missing_by_patient)) {
  if ("Troponin_high" %in% colnames(missing_by_patient)) {
    missing_by_patient$Troponin = missing_by_patient$Troponin_normal +   missing_by_patient$Troponin_high
  }else{
  colnames(missing_by_patient)[which(colnames(missing_by_patient)=="Troponin_normal")] <- "Troponin"
  }
}

if ("DDU" %in% colnames(missing_by_patient)){
  if ("FEU" %in% colnames(missing_by_patient)) {
    ct = ct + 1
  }else{
  colnames(missing_by_patient)[which(colnames(missing_by_patient)=="DDU")] <- "D_Dimer"
  }
}

if ("FEU" %in% colnames(missing_by_patient)) {
  if ("D-Dimer" %in% colnames(missing_by_patient)) {
    missing_by_patient$D_Dimer = missing_by_patient$FEU +   missing_by_patient$D_Dimer
  }else{
  colnames(missing_by_patient)[which(colnames(missing_by_patient)=="FEU")] <- "D_Dimer"
  }
}


drops = c("Troponin_high","Troponin_normal","FEU","DDU")
missing_by_patient= missing_by_patient[,!(names(missing_by_patient) %in% drops)]

allLabs <- c("Fibrinogen","Procalcitonin","D_Dimer","Troponin","CRP","Ferritin","LDH","Lymphocyte","PT","Albumin","Neutrophil","AST","ALT","Bilirubin","Leukocytes","Creatinine")
labsPresent <- allLabs[ allLabs %in% colnames(missing_by_patient)]

lab_names = labsPresent

#Obtain total number of observations per lab for TE patients

uniq_vals <-
  apply(missing_by_patient, 2, function(x) {
    length(unique(x))
  })


#Remove patients that have no lab values

missing_by_patient <- missing_by_patient[, uniq_vals > 1] %>%
  mutate(sum_labs = rowSums(across(all_of(lab_names)))) %>% 
  filter(sum_labs > 0)

#Reduce dataframe to matrix with labs and values

x_mat <- missing_by_patient[, allLabs]
```

```{r}
# set.seed(1)
# m <- topicmodels::LDA(x_mat, method = "Gibbs", k = 8,  control = list(alpha = 0.1))
```

```{r cache=TRUE}
exclusivity <- function(mod.out, M = 10, frexw = .7) {
  w <- frexw
  if (length(mod.out$beta$logbeta) != 1) stop("Exclusivity calculation only designed for models without content covariates")
  tbeta <- t(exp(mod.out$beta$logbeta[[1]]))
  s <- rowSums(tbeta)
  mat <- tbeta / s # normed by columns of beta now.

  ex <- apply(mat, 2, rank) / nrow(mat)
  fr <- apply(tbeta, 2, rank) / nrow(mat)
  frex <- 1 / (w / ex + (1 - w) / fr)
  index <- apply(tbeta, 2, order, decreasing = TRUE)[1:M, ]
  out <- vector(length = ncol(tbeta))
  for (i in 1:ncol(frex)) {
    out[i] <- sum(frex[index[, i], i])
  }
  out
}

x_dfm <- x_mat %>%
  rownames_to_column("id") %>%
  pivot_longer(-id, names_to = "lab", values_to = "n") %>%
  tidytext::cast_dfm(id, lab, n)
```


```{r cache=TRUE}
system.time(many_models <- data.frame(K = seq(2, 8, 1)) %>%
  mutate(topic_model = furrr::future_map(K, ~ stm(
    x_dfm,
    K = .,
    seed = TRUE,
    verbose = FALSE
  ))))

heldout <- make.heldout(x_dfm)

k_result <- many_models %>%
  mutate(
    exclusivity = map(topic_model, exclusivity),
    semantic_coherence = map(topic_model, semanticCoherence, x_dfm),
    eval_heldout = map(topic_model, eval.heldout, heldout$missing),
    residual = map(topic_model, checkResiduals, x_dfm),
    bound = map_dbl(topic_model, function(x) max(x$convergence$bound)),
    lfact = map_dbl(topic_model, function(x) lfactorial(x$settings$dim$K)),
    lbound = bound + lfact,
    iterations = map_dbl(topic_model, function(x) length(x$convergence$bound))
  )

topic_diagnostics = k_result %>%
  transmute(K,
    `Lower bound` = lbound,
    Residuals = map_dbl(residual, "dispersion"),
    `Semantic coherence` = map_dbl(semantic_coherence, mean),
    `Held-out likelihood` = map_dbl(eval_heldout, "expected.heldout")
  ) %>%
  gather(Metric, Value, -K) 


topic_diagnostics_wide = data.frame(matrix(ncol=0,nrow=7))
topic_diagnostics_wide$K= topic_diagnostics$K[1:7]
topic_diagnostics_wide$LB= topic_diagnostics$Value[1:7]
topic_diagnostics_wide$Res = topic_diagnostics$Value[8:14]
topic_diagnostics_wide$SC = topic_diagnostics$Value[15:21]
topic_diagnostics_wide$HL = topic_diagnostics$Value[22:28]

ind = c(which.max(topic_diagnostics_wide$LB)+1,which.max(topic_diagnostics_wide$SC)+1,which.max(topic_diagnostics_wide$HL)+1,which.min(topic_diagnostics_wide$Res)+1)

Modes <- function(x) {
  ux <- unique(x)
  tab <- tabulate(match(x, ux))
  ux[tab == max(tab)]
}

Modes(ind)

        
                                                                      topic_diagnostics_plot = k_result %>%
  transmute(K,
    `Lower bound` = lbound,
    Residuals = map_dbl(residual, "dispersion"),
    `Semantic coherence` = map_dbl(semantic_coherence, mean),
    `Held-out likelihood` = map_dbl(eval_heldout, "expected.heldout")
  ) %>%
  gather(Metric, Value, -K) %>%
  ggplot(aes(K, Value, color = Metric)) +
  geom_line() +
  geom_point() +
  guides(color = FALSE) +
  scale_x_continuous(breaks = seq(2, 12, 2)) +
  facet_wrap(~Metric, scales = "free_y") +
  labs(
    x = "K (number of topics)",
    y = NULL,
    title = "Model diagnostics by number of topics"
  )

topic_diagnostics_plot




```

We want to: maximize likelihood, lower bound, semantic coherence and minimize residuals.
- https://scholar.harvard.edu/files/dtingley/files/multimod.pdf
- http://proceedings.mlr.press/v22/taddy12/taddy12.pdf

```{r}

#Estimate the structural topic model using semi-collapsed variational EM. 

#Change K depending on what the optimal number of topics is based on the plots above (maximize likelihood, lower bound, semantic coherence, and minimize residuals)




K <- Modes(ind)[1]
stmfit <- stm(x_dfm, K = K, verbose = FALSE, init.type = "Spectral", seed = TRUE)
stmfit_beta <- stmfit$beta



```

```{r}

topicQuality = topicQuality(stmfit, documents = x_dfm)
```

The darker the colors, the more the presence of that lab values contribute to that topic.
```{r}

#Beta matrix contains log probabilities of labs in topics. Generate heat map of beta values for each lab.


K=Modes(ind)[1]
beta_mat <- exp(stmfit$beta$logbeta[[1]])
colnames(beta_mat) <- stmfit$vocab
rownames(beta_mat) <- paste("Topic", 1:K)
heatmap((beta_mat), main = "Ordered",
        col= colorRampPalette(brewer.pal(8, "Blues"))(25))
legend(x="topleft", legend=c("min", "ave", "max"), 
     fill=colorRampPalette(brewer.pal(8, "Blues"))(3))

labs = stmfit$vocab

```

```{r}
theta <- stmfit$theta

corrFunc <- function(var1, var2) {
  result = cor.test(var1,var2)
  data.frame(var1, var2, result[c("estimate","p.value","statistic","method")], 
             stringsAsFactors=FALSE)
}

## Pairs of variables for which we want correlations

theta <- stmfit$theta
K = Modes(ind)[1]

TE_results <- lapply(1:K, function(x) cor.test(theta[, x], as.numeric(missing_by_patient$TE))[c("estimate","p.value","statistic","method")])

Severity_results <- lapply(1:K, function(x) cor.test(theta[, x], as.numeric(missing_by_patient$severe))[c("estimate","p.value","statistic","method")])

Neuro_results <- lapply(1:K, function(x) cor.test(theta[, x], as.numeric(missing_by_patient$Neuro))[c("estimate","p.value","statistic","method")])

ARDs_results <- lapply(1:K, function(x) cor.test(theta[, x], as.numeric(missing_by_patient$ARDs))[c("estimate","p.value","statistic","method")])


```

```{r}

theta <- stmfit$theta
K = Modes(ind)

#TE 

topic_df <- data.frame(theta, te = missing_by_patient$TE) %>% 
  pivot_longer(- te, names_to = 'topic') %>% 
  mutate(topic = gsub('X', 'Topic ', topic))

TE_boxplot <- topic_df %>% 
  ggplot(aes(x = te, y = value)) +
  labs(
    x = "Thrombotic Event",
    y = "Topic Value"
  ) +
  geom_boxplot(alpha = 0.2) +
  facet_wrap(~ topic, scales = 'free') +
  NULL

TE_boxplot


#Severity

topic_df <- data.frame(theta, Severity = missing_by_patient$severe) %>% 
  pivot_longer(- Severity, names_to = 'topic') %>% 
  mutate(topic = gsub('X', 'Topic ', topic))

Severity_boxplot <- topic_df %>% 
  ggplot(aes(x = Severity, y = value)) +
  labs(
    x = "Severity",
    y = "Topic Value"
  ) +
  geom_boxplot(alpha = 0.2) +
  facet_wrap(~ topic, scales = 'free') +
  NULL

Severity_boxplot

#Neuro

topic_df <- data.frame(theta, Neuro = missing_by_patient$Neuro) %>% 
  pivot_longer(- Neuro, names_to = 'topic') %>% 
  mutate(topic = gsub('X', 'Topic ', topic))

Neuro_boxplot <- topic_df %>% 
  ggplot(aes(x = Neuro, y = value)) +
  labs(
    x = "Neurological Event",
    y = "Topic Value"
  ) +
  geom_boxplot(alpha = 0.2) +
  facet_wrap(~ topic, scales = 'free') +
  NULL

Neuro_boxplot

#ARDs

topic_df <- data.frame(theta, ARDs = missing_by_patient$ARDs) %>% 
  pivot_longer(- ARDs, names_to = 'topic') %>% 
  mutate(topic = gsub('X', 'Topic ', topic))

ARDs_boxplot <- topic_df %>% 
  ggplot(aes(x = ARDs, y = value)) +
  labs(
    x = "Acute Respiratory Distress",
    y = "Topic Value"
  ) +
  geom_boxplot(alpha = 0.2) +
  facet_wrap(~ topic, scales = 'free') +
  NULL

ARDs_boxplot

```
```{r}
#Apply NWU topics to new data:

make_matrix <- function(df,rownames = NULL){
  my_matrix <-  as.matrix(df)
  if(!is.null(rownames))
    rownames(my_matrix) = rownames
  my_matrix
}


new_mat = x_mat[,order(colnames(x_mat))]

x_matrix = make_matrix(new_mat)
props = apply_betas %*% t(x_matrix)
new_props = t(props)
actual_props = t(props)
for (i in 1:nrow(new_props)){
  actual_props[i,] = prop.table(new_props[i,])
}
theta = actual_props

K = 6

TE_results_comb<- lapply(1:K, function(x) cor.test(theta[, x], as.numeric(missing_by_patient$TE))[c("estimate","p.value","statistic","method")])

Severity_results_comb <- lapply(1:K, function(x) cor.test(theta[, x], as.numeric(missing_by_patient$severe))[c("estimate","p.value","statistic","method")])

Neuro_results_comb <- lapply(1:K, function(x) cor.test(theta[, x], as.numeric(missing_by_patient$Neuro))[c("estimate","p.value","statistic","method")])

ARDs_results_comb <- lapply(1:K, function(x) cor.test(theta[, x], as.numeric(missing_by_patient$ARDs))[c("estimate","p.value","statistic","method")])

#TE 

topic_df <- data.frame(theta, te = missing_by_patient$TE) %>% 
  pivot_longer(- te, names_to = 'topic') %>% 
  mutate(topic = gsub('X', 'Topic ', topic))

TE_boxplot_comb <- topic_df %>% 
  ggplot(aes(x = te, y = value)) +
  labs(
    x = "Thrombotic Event",
    y = "Topic Value"
  ) +
  geom_boxplot(alpha = 0.2) +
  facet_wrap(~ topic, scales = 'free') +
  NULL

TE_boxplot_comb


#Severity

topic_df <- data.frame(theta, Severity = missing_by_patient$severe) %>% 
  pivot_longer(- Severity, names_to = 'topic') %>% 
  mutate(topic = gsub('X', 'Topic ', topic))

Severity_boxplot_comb <- topic_df %>% 
  ggplot(aes(x = Severity, y = value)) +
  labs(
    x = "Severity",
    y = "Topic Value"
  ) +
  geom_boxplot(alpha = 0.2) +
  facet_wrap(~ topic, scales = 'free') +
  NULL

Severity_boxplot_comb

#Neuro

topic_df <- data.frame(theta, Neuro = missing_by_patient$Neuro) %>% 
  pivot_longer(- Neuro, names_to = 'topic') %>% 
  mutate(topic = gsub('X', 'Topic ', topic))

Neuro_boxplot_comb <- topic_df %>% 
  ggplot(aes(x = Neuro, y = value)) +
  labs(
    x = "Neurological Event",
    y = "Topic Value"
  ) +
  geom_boxplot(alpha = 0.2) +
  facet_wrap(~ topic, scales = 'free') +
  NULL

Neuro_boxplot_comb

#ARDs

topic_df <- data.frame(theta, ARDs = missing_by_patient$ARDs) %>% 
  pivot_longer(- ARDs, names_to = 'topic') %>% 
  mutate(topic = gsub('X', 'Topic ', topic))

ARDs_boxplot_comb <- topic_df %>% 
  ggplot(aes(x = ARDs, y = value)) +
  labs(
    x = "Acute Respiratory Distress",
    y = "Topic Value"
  ) +
  geom_boxplot(alpha = 0.2) +
  facet_wrap(~ topic, scales = 'free') +
  NULL

ARDs_boxplot_comb


```






To get the interactive visualization, please uncomment the following lines,
Run All (not Knit).
Please share a screenshot of the result if possible!
You may have to install the tm package with `install.packages('tm')`.
```{r}
save(topic_diagnostics,topicQuality,TE_results,Severity_results,Neuro_results,ARDs_results,TE_results_comb,Severity_results_comb,Neuro_results_comb,ARDs_results_comb,stmfit_beta,beta_mat,labs,
  file = paste("results",params$results_file,sep=""))

ggsave('figs/topic_diagnostics_plot.png', topic_diagnostics_plot)
ggsave('figs/TE_boxplot.png', TE_boxplot)
ggsave('figs/Severity_boxplot.png', Severity_boxplot)
ggsave('figs/Neuro_boxplot.png', Neuro_boxplot)
ggsave('figs/ARDs_boxplot.png', ARDs_boxplot)
ggsave('figs/TE_boxplot_comb.png', TE_boxplot_comb)
ggsave('figs/Severity_boxplot_comb.png', Severity_boxplot_comb)
ggsave('figs/Neuro_boxplot_comb.png', Neuro_boxplot_comb)
ggsave('figs/ARDs_boxplot_comb.png', ARDs_boxplot_comb)








```

```{r}
beepr::beep()
```

```{r}
knitr::knit_exit()
```


