---
title: "Missingness analysis"
author: Amelia Tan, Arianna Dagliati, Trang Le
date: "27 October 2020"
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "htmls") })
---
```{r setup}
source("R/utils.R")
required_pkgs <- c("tidyverse", "lubridate", "ggridges", "cowplot", "naniar", "DT","RColorBrewer")
install_required(required_pkgs)

library(tidyverse)
library(DT)
library(lubridate)
library(naniar)
library(cowplot)
library(ggridges)
library(RColorBrewer)

theme_set(theme_bw())
theme_update(
  legend.title = element_blank(),
  panel.grid.minor = element_blank()
)
```


```{r warning=FALSE, message=FALSE}

#Read in local patient summary

data_dir <- "../4ce/Input/"
demo_raw <-
  readr::read_csv(
    file.path(data_dir, "LocalPatientSummary.csv"),
    col_types = list(patient_num = readr::col_character()),
    na = "1900-01-01"
  ) %>%
  mutate(across(ends_with("_date") & where(is.character), lubridate::mdy)) %>%
  mutate(
    last_discharge_date = if_else(
      !is.na(death_date) & death_date < last_discharge_date,
      death_date,
      last_discharge_date
    )
  )

#Read in thrombotic icd codes. Add column "truncated code" for first three characters of each code. Extract unique truncated codes. 

thrombo_codes <- read_csv("https://raw.githubusercontent.com/covidclinical/Phase2.1AKIRPackage/ac19716a4586f45c398728fcd821ca9d5baffe45/FourCePhase2.1AKI/data-raw/thromb_icd_code.csv") %>%
  mutate(truncated_code = substr(icd_code, 1, 3)) %>%
  pull(truncated_code) %>%
  unique()

#Extract information for patients with codes for thrombotic events

patient_obs <- read_csv(
  file.path(data_dir, "LocalPatientObservations.csv"),
  col_types = list(patient_num = readr::col_character()),
  na = "-999"
)

te_patients <- patient_obs %>%
  filter(
    concept_type == "DIAG-ICD10",
    concept_code %in% thrombo_codes,
    days_since_admission >= 0
  )
```

## Read in files 
and convert to wide format

```{R message=FALSE}
# 4CE long format Labs Data


# Code, Descriptions and Ranges
lab_mapping <- read_csv("public-data/loinc-map.csv")
lab_bounds <- read_csv("public-data/lab_bounds.csv")
lab_names <- lab_bounds$short_name

# load('public-data/code.dict.rda')
```


```{R}
# setdiff(patient_obs$concept_code %>% unique(), lab_bounds$LOINC)

#Generate a table that contains information about each lab for each individual patient. Clean up data by removing observations with missing lab names or < 0 days since admission. Each observation is identified by their patient number and days since admission, and each lab name and value is included. If there is more than one lab value for that identifier, they are averaged. Also add severity column. 

patient_obs_wide <- patient_obs %>%
  left_join(lab_bounds, by = c("concept_code" = "LOINC")) %>%
  filter(days_since_admission >= 0, !is.na(short_name)) %>%
  select(-concept_code) %>%
  pivot_wider(
    id_cols = c(patient_num, days_since_admission),
    names_from = short_name,
    values_from = value,
    values_fn = mean
  ) %>%
  left_join(select(demo_raw, patient_num, severe), by = "patient_num") %>%
  mutate(severity = factor(severe) %>%
    fct_recode("severe" = "1", "nonsevere" = "0")) %>%
  select(-severe)

#Extract unique lab names

lab_names <- intersect(lab_bounds$short_name, names(patient_obs_wide))
```


```{R}

#Extract observations from patient_obs_wide that have days since admission between 0 and 20 and a thrombotic event code. Remove columns for patient ID and days since admisison.

te_labs <- patient_obs_wide %>%
  filter(
    days_since_admission <= 20,
    days_since_admission >= 0
  ) %>%
  mutate(te = patient_num %in% unique(te_patients$patient_num)) %>%
  select(-c(patient_num, days_since_admission))

#Generate a ggplot of the missingness inside a dataframe. Cells are colored according to missingness (black indicates a missing cell and grey indicates a present cell). 

vis_miss(te_labs, sort_miss = TRUE)
```

```{R}

#Calculate the Spearman correlation between lab missingness indicators to see if certain labs tend to be missing together. First, isolate only the first 20 medical visits for each patient. Then, for each time point, calculate the spearman correlation matrix for the missingness indicators for each lab. Generate a list of these matrices, and for each lab combination, see if at least 70% of the spearman correlations from each time point are > 0.5. If they are, isolate the combination. For all combinations that pass the threshold, generate a boxplot of the correlation values from each time point. 


corr_miss <- function(dat, min_corr, prop_greater) {

  te_labs2 <- dat %>% 
    filter(days_since_admission <= 20,
           days_since_admission >= 0) %>% 
    mutate(te = patient_num %in% unique(te_patients$patient_num))
  
  sim_mat = matrix(rep(0),ncol(te_labs2)-4,ncol(te_labs2)-4)
  mat_list = list()
  
  for (h in 0:max(te_labs2$days_since_admission)){
  
  data1 = te_labs2[which(te_labs2$days_since_admission==h),]
  
  na_stats <- data1 %>% 
    select(- c(days_since_admission,patient_num,te,severity)) %>% 
    is.na() %>% 
    `!` 
  
  #trying something
  for (i in 1:ncol(na_stats)){
    na_stats[,i] <- as.integer(as.logical(na_stats[,i]))
  }
  
  sim_mat = matrix(rep(0),ncol(na_stats),ncol(na_stats))
  
  for (i in 1:ncol(na_stats)){
    for (j in 1:ncol(na_stats)){
      sim_mat[i,j] <- cor(na_stats[,i],na_stats[,j],method="spearman")
    }
  }
  
  rownames(sim_mat) <- colnames(na_stats)
  colnames(sim_mat) <- colnames(na_stats)
  
  mat_list[[h+1]] = sim_mat
  
  }
  
  df <- data.frame(matrix(ncol=0,nrow=21))
  names = list()
  final_names = c()
  count = 0
  for (i in 1:ncol(na_stats)){
    for (j in 1:ncol(na_stats)){
      name1 <- c(row.names(mat_list[[1]])[i],colnames(mat_list[[1]])[j])
      name1 <- do.call(paste, c(as.list(name1), sep = ", "))
      name2 <- c(row.names(mat_list[[1]])[j],colnames(mat_list[[1]])[i])
      name2 <- do.call(paste, c(as.list(name2), sep = ", "))
      if (i != j && !(name1 %in% final_names) && !(name2 %in% final_names)) {
        test_vec = c()
        for (h in 1:length(mat_list)){
          test_vec[h] = mat_list[[h]][i,j]
        }
        if (mean(abs(test_vec) > min_corr) > prop_greater) {
          count = count+1
          final_names[count] = name1
          df[,count] = test_vec
        }
  
      }
      }
    }
  
  colnames(df) <- final_names
  
  results <- list()
  results[[1]] <- df
  results[[2]] <- mat_list
  
  return(results)

}

results = corr_miss(patient_obs_wide,0.5,0.7)

boxplot(results[[1]],ylim=c(0, 1),xlab='Lab Combination',ylab='Spearman Correlation',main='Correlation Between Lab Missing Indicators')




#Generate a heat map for the Spearman correlation between lab missingness at a given time point (in this example, second time point). Legend displays the range of Spearman correlation values. 

which_mat = 2

heatmap(results[[2]][[which_mat]],scale="none",,cexRow=1,cexCol=1,margins=c(8,6))

legend(x="bottomright", legend=c(round(min(results[[2]][[which_mat]]),2),round(((max(results[[2]][[which_mat]])+min(results[[2]][[which_mat]]))/2),2), max(results[[2]][[which_mat]])), 
     fill=colorRampPalette(brewer.pal(8, "Oranges"))(3))

```


```{r}

#Plot pattern of missingness using upset plot. Visualize number of missing values for sets of data. Uses 7 sets (7 variables) and 10 intersections (10 interactions)

gg_miss_upset(te_labs, nsets = 7, nintersects = 10)

#Plot number of missings for each lab, broken down by no thrombotic event vs. presence of thrombotic event(s)

gg_miss_fct(x = te_labs, fct = te)

#Plot number of missing lab values for each observation (patient time point)

gg_miss_case(te_labs)

#Plot number of missing values for each lab

gg_miss_var(te_labs) +
  theme(
    panel.grid.minor = element_blank(),
    panel.grid.major.y = element_blank()
  ) +
  scale_y_continuous(expand = expansion(0, c(0, 1000))) +
  xlab(NULL)

#Plot percent of missing values for each lab

gg_miss_var(te_labs, show_pct = TRUE) +
  theme(
    panel.grid.minor = element_blank(),
    panel.grid.major.y = element_blank()
  ) +
  scale_y_continuous(expand = expansion(0, c(0, 1))) +
  xlab(NULL)
# mcar_test(te_labs)

# te_labs %>%
#   finalfit::missing_pairs(dependent = "te",
#                           explanatory = lab_names,
#                           position = "fill")
```


```{R}
# check NAs in the Wide format

#Create True / False dataframe for whether or not a lab is missing for each observation

na_stats <- patient_obs_wide %>%
  select(-c(patient_num, days_since_admission)) %>%
  is.na() %>%
  `!`()

#Create new dataframe containing number of values and proportion of values existing per lab 

na_df <- data.frame(
  value_existed = colSums(na_stats),
  prop_existed = colMeans(na_stats)
) %>%
  rownames_to_column("lab") %>%
  mutate(
    prop_na = 1 - prop_existed,
    lab = fct_reorder(lab, value_existed)
  )

#Plot number of values per lab

n_values <- na_df %>%
  ggplot(aes(x = value_existed, y = lab)) +
  geom_col() +
  labs(x = "Number of values", y = NULL)

#Plot proportion of values per lab. Valid value indicates existing values. 

na_prob <- na_df %>%
  rename("Valid value" = prop_existed, "NA" = prop_na) %>%
  pivot_longer(c(`Valid value`, `NA`)) %>%
  ggplot(aes(x = value, y = lab, fill = name)) +
  geom_col() +
  scale_fill_discrete(guide = guide_legend(reverse = TRUE)) +
  labs(x = "Proportion", y = NULL) +
  # guides(fill = guide_legend(reverse = TRUE))
  theme(
    axis.text.y = element_blank(),
    legend.key.width = unit(6, "mm"),
    legend.key.height = unit(4, "mm"),
    legend.position = "bottom"
  )

plot_grid(n_values, na_prob, nrow = 1, axis = "b", align = "h")

penn_na_df <- na_df
```


```{r}

#Keep the patient num and days_since_admission variables, and convert the individual lab name variables into one single variable where each observation is the name of the lab, and create another variable for the corresponding lab values. 

patient_obs_long <- patient_obs_wide %>%
  pivot_longer(-c(patient_num, days_since_admission, severity),
    names_to = "lab", values_to = "value",
    values_drop_na = TRUE
  )
```


```{r fig.show='hide'}
# number of patients per lab value

#For an individual lab, obtain number of lab observations per patient. Plot a histogram of the number of lab observations for each patient. Create dataframe containing counts and density for each lab.

get_pat_labs <- function(labi) {
  patients_labs <- patient_obs_long %>%
    filter(lab == labi) %>%
    count(patient_num) %>%
    pull(n) %>%
    hist(., breaks = seq(0, max(.), 1), main = labi)

  patients_labs[[1]] <- patients_labs[[1]][-1]
  data.frame(
    do.call(cbind.data.frame, patients_labs[1:3]),
    lab = labi
  )
}

#Run function for each lab and bind the rows from each dataframe

patient_lab <- lapply(unique(patient_obs_long$lab), get_pat_labs) %>%
  bind_rows()

penn_patient_lab <- patient_lab


#Replicate each row by the length of the dataframe

melt_patient_lab <- patient_lab[rep(1:nrow(patient_lab), patient_lab$counts), ]
```

```{r}
melt_patient_lab %>%
  filter(breaks < 30) %>%
  mutate(lab = fct_infreq(lab)) %>%
  ggplot(aes(x = breaks, y = lab, fill = lab, height = ..count..)) +
  # geom_density_ridges(stat = "identity", scale = 2) +
  geom_ridgeline(stat = "binline", binwidth = 1, scale = 0.001) +
  # scale_color_viridis_d(option = 'C') +
  scale_fill_viridis_d(option = "C", guide = FALSE) +
  labs(y = NULL) +
  # coord_flip() +
  # facet_wrap(~ lab, scales = 'free') +
  NULL
```


## Number of observation (days) per patient
```{R}
days_count_min_max <-
  # multiple_imps %>%
  patient_obs_wide %>%
  group_by(patient_num, severity) %>%
  summarise(
    n_values = n_distinct(days_since_admission),
    min_day = min(days_since_admission),
    max_day = max(days_since_admission),
    .groups = "drop"
  ) %>%
  # left_join(demo_raw, by = "patient_num") %>%
  # mutate(time_obs = max_day - min_day,
  #        severity = factor(severe)) %>%
  # select(-patient_num) %>%
  add_count(severity, name = "n_severity") %>%
  {
    .
  }

penn_agg_n_values <- days_count_min_max %>%
  count(severity, n_severity, n_values,
    name = "n_nvals"
  )
penn_agg_max_day <- days_count_min_max %>%
  count(severity, n_severity, max_day,
    name = "n_maxday"
  )

penn_agg_n_values %>%
  ggplot(aes(x = n_nvals)) +
  geom_bar()

(n_severe <- sum(days_count_min_max$severity == "severe"))
(n_nonsevere <- sum(days_count_min_max$severity == "nonsevere"))
```

## Histogram of the number of days with at least one observation
```{R}
penn_agg_n_values %>%
  ggplot(aes(x = n_values, y = n_nvals, fill = severity)) +
  geom_col(alpha = 0.5) +
  scale_fill_brewer(palette = "Dark2", direction = -1) +
  labs(
    fill = "Severe?",
    x = "Number of days with data",
    y = "Count"
  )
```

## Histogram of length of stay
i.e. last day with observation-first day with observation

We need to check for readmission here.
```{R}
penn_agg_max_day %>%
  ggplot(aes(x = max_day, y = n_maxday, fill = severity)) +
  geom_col(alpha = 0.5) +
  scale_fill_brewer(palette = "Dark2", direction = -1) +
  labs(
    fill = "Severe?",
    x = "Number of days with data",
    y = "Count"
  )
```

## Analyze missingness and frequency of measures for each lab

```{r}

#Count the number of observations for each individual patient.  Then count the number of patients that have x severe observations and z nonsevere observations for a given lab (for the range of x and z observations for a given lab). Add a column for both severities. Add another column for the proportion of severe patients that have x severe observations for a given lab out of all severe patients. Do this for non-severe and both severities. 

per_lab <- patient_obs_long %>%
  group_by(lab, patient_num, severity) %>%
  count(name = "n_obs") %>%
  ungroup() %>%
  group_by(lab, severity) %>%
  count(n_obs) %>%
  ungroup() %>%
  pivot_wider(names_from = severity, values_from = n, values_fill = 0) %>%
  mutate(both_severities = nonsevere + severe) %>%
  mutate(
    prop_nonsevere = nonsevere / n_nonsevere,
    prop_severe = severe / n_severe,
    prop_both = both_severities / nrow(days_count_min_max)
  )

lab_medians <-
  patient_obs_long %>%
  add_count(lab, name = "total_obs") %>%
  group_by(lab) %>%
  mutate(total_patients = length(unique(patient_num))) %>%
  add_count(patient_num, name = "n_obs_patients") %>%
  mutate(
    median_obs_per_patient = median(n_obs_patients),
    n_greater0 = n_obs_patients > median_obs_per_patient,
    n_greater1 = n_obs_patients > median_obs_per_patient + 1,
    n_greater2 = n_obs_patients > median_obs_per_patient + 2
  ) %>%
  group_by(severity) %>%
  mutate(each_med_obs_per_patient = median(n_obs_patients)) %>%
  ungroup(severity) %>%
  select(-c(days_since_admission, value)) %>%
  distinct() %>%
  group_by(lab) %>%
  mutate(across(contains("n_greater"), sum)) %>%
  select(-c(n_obs_patients, patient_num)) %>%
  distinct() %>%
  pivot_wider(names_from = severity, values_from = each_med_obs_per_patient) %>%
  rename(
    "median_obs_per_severe_patient" = severe,
    "median_obs_per_non_severe_patient" = nonsevere
  ) %>%
  select(lab, total_obs, total_patients, starts_with("med"), starts_with("n_"))

lab_medians %>%
  datatable(rownames = FALSE)
```


`n_greater0` shows the number of patients who had more observations than the median.
`n_greater1` shows the number of patients who had more observations than the median + 1.
`n_greater2` shows the number of patients who had more observations than the median + 2.

```{r eval=FALSE, include=FALSE}
lab_medians %>%
  select(lab, total_obs, total_patients) %>%
  pivot_longer(-lab) %>%
  ggplot(aes(x = value, y = fct_reorder(lab, value))) +
  geom_col() +
  facet_grid(cols = vars(name), scales = "free_x", space = "free") +
  labs(y = NULL)
```

In the figure below:

- Grey dash line: `Reference Low`
- Grey solid line: `Reference High`
- Black dash line: `lower bound outlier` (QC)
- Black solid line: `upper bound outlier` (QC)

```{r fig.width=9, fig.height=11}
patient_obs_long %>%
  left_join(lab_bounds, by = c("lab" = "short_name")) %>%
  ggplot(aes(y = severity, x = value, fill = severity)) +
  geom_violin() +
  scale_fill_brewer(palette = "Dark2", guide = guide_legend(reverse = TRUE)) +
  labs(y = NULL, x = NULL) +
  geom_vline(aes(xintercept = `Reference Low`), linetype = "dashed", color = "grey") +
  geom_vline(aes(xintercept = `Reference High`), color = "grey") +
  geom_vline(aes(xintercept = LB), linetype = "dashed") +
  geom_vline(aes(xintercept = UB)) +
  facet_wrap(~lab, scales = "free", ncol = 2, strip.position = "left") +
  theme(axis.text.y = element_blank())
```

## Missing data heatmap

"Binned" heatmap

```{r fig.width=12}

#Generate a heatmap in which darker purple colors indicate a higher proportion of patients in a given bin (detailing the number of values a patient has for a given lab). 

per_lab %>%
  mutate(obs_bin = cut(
    n_obs,
    breaks = c(0:15, 20, 30, max(n_obs))
  )) %>%
  group_by(lab, obs_bin) %>%
  summarise(
    both_severities = sum(both_severities),
    severe = sum(severe),
    nonsevere = sum(nonsevere),
    .groups = "drop"
  ) %>%
  select(lab, obs_bin, both_severities, severe, nonsevere) %>%
  pivot_longer(c(both_severities, severe, nonsevere)) %>%
  mutate(name = name %>% fct_recode(
    "All patients" = "both_severities",
    "Severe patients" = "severe",
    "Non-severe patients" = "nonsevere"
  )) %>%
  ggplot(aes(x = obs_bin, fill = value, y = fct_reorder(lab, value))) +
  geom_tile(colour = "white", size = 0.2) +
  geom_text(aes(label = value), colour = "white", size = 2) +
  scale_y_discrete(expand = c(0, 0)) +
  scale_fill_gradient(low = "lightgrey", high = "darkblue") +
  facet_wrap(~name, nrow = 1) +
  labs(
    x = "Binned number of values a patient has for each lab",
    y = NULL, fill = "# patients"
  ) +
  theme(
    panel.grid.major = element_blank(),
    legend.position = c(0.93, 0.2),
    axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5),
    axis.ticks.y = element_blank()
  )
```


```{r fig.width=12}
per_lab %>%
  mutate(obs_bin = cut(n_obs, breaks = c(0:15, 20, 30, max(n_obs)))) %>%
  group_by(lab, obs_bin) %>%
  summarise(
    prop_both = sum(prop_both, na.rm = TRUE),
    prop_severe = sum(prop_severe, na.rm = TRUE),
    prop_nonsevere = sum(prop_nonsevere, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  select(lab, obs_bin, prop_both, prop_severe, prop_nonsevere) %>%
  pivot_longer(c(prop_both, prop_severe, prop_nonsevere)) %>%
  mutate(name = name %>% fct_recode(
    "Compared to all patients" = "prop_both",
    "Compared to all severe patients" = "prop_severe",
    "Compared to all non-severe patients" = "prop_nonsevere"
  )) %>%
  ggplot(aes(x = obs_bin, fill = value, y = fct_reorder(lab, value))) +
  geom_tile(colour = "white", size = 0.2) +
  geom_text(aes(label = round(value, 2)), colour = "white", size = 2) +
  scale_y_discrete(expand = c(0, 0)) +
  scale_fill_gradient(
    low = "lightgrey", high = "darkblue",
    labels = scales::percent_format(accuracy = 1L)
  ) +
  facet_wrap(~name, nrow = 1) +
  labs(
    x = "Binned number of values a patient has for each lab",
    y = NULL, fill = "% patients"
  ) +
  theme(
    panel.grid.major = element_blank(),
    legend.position = c(0.93, 0.2),
    axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5),
    axis.ticks.y = element_blank()
  )
```

Denominator: total number of patients, total number of non-severe patients,
and total number of severe patients, respectively.

```{r}
per_lab %>%
  select(lab, n_obs, severe, nonsevere) %>%
  filter(n_obs <= 90) %>%
  pivot_longer(c(severe, nonsevere)) %>%
  mutate(
    lab = fct_reorder(lab, n_obs),
    name = name %>% fct_recode(
      "Severe patients" = "severe",
      "Non-severe patients" = "nonsevere"
    )
  ) %>%
  ggplot(aes(x = n_obs, fill = name, y = value)) +
  geom_col() +
  scale_fill_brewer(palette = "Dark2", direction = -1) +
  facet_wrap(~lab, scales = "free") +
  labs(
    x = "Number of values a patient has for each lab",
    y = NULL, fill = "# patients"
  ) +
  theme(
    panel.grid.major = element_blank(),
    legend.position = c(0.9, 0.2),
    axis.ticks.y = element_blank()
  )
```


```{r}
sessionInfo()
```

```{r}
save(penn_na_df, penn_patient_lab, penn_agg_n_values, penn_agg_max_day, 
  file = "results/penn-results.Rdata"
)
```
