---
title: "Missingness analysis"
author: Amelia Tan, Arianna Dagliati, Trang Le
date: "27 October 2020"
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "htmls") })
---

```{r setup, warning=FALSE, message=FALSE}
library(tidyverse)
library(DT)
library(lubridate)
library(naniar)
library(cowplot)
library(ggridges)

theme_set(theme_bw())
theme_update(legend.title = element_blank(),
             panel.grid.minor = element_blank())
```


```{r warning=FALSE, message=FALSE}
data_dir <- '../4ce/Input/'
demo_raw <-
  readr::read_csv(
    file.path(data_dir, "LocalPatientSummary.csv"),
    col_types = list(patient_num = readr::col_character()),
    na = "1900-01-01"
  ) %>%
  mutate(across(ends_with("_date") & where(is.character), lubridate::mdy)) %>%
  mutate(
    last_discharge_date = if_else(
      !is.na(death_date) & death_date < last_discharge_date,
      death_date,
      last_discharge_date
    )
  )
thrombo_codes <- read_csv("https://raw.githubusercontent.com/covidclinical/Phase2.1AKIRPackage/ac19716a4586f45c398728fcd821ca9d5baffe45/FourCePhase2.1AKI/data-raw/thromb_icd_code.csv") %>%
  mutate(truncated_code = substr(icd_code, 1, 3)) %>%
  pull(truncated_code) %>%
  unique()

patient_obs <- read_csv(
  file.path(data_dir, "LocalPatientObservations.csv"),
  col_types = list(patient_num = readr::col_character()),
  na = "-999"
)

te_patients <- patient_obs %>%
  filter(
    concept_type == "DIAG-ICD10",
    concept_code %in% thrombo_codes,
    days_since_admission >= 0
  ) 
```

## Read in files 
and convert to wide format

```{R message=FALSE}
#4CE long format Labs Data


#Code, Descriptions and Ranges
lab_mapping <- read_csv('public-data/loinc-map.csv')
lab_bounds <- read_csv('public-data/lab_bounds.csv')
lab_names <- lab_bounds$short_name

# load('public-data/code.dict.rda')
```


```{R}
# setdiff(patient_obs$concept_code %>% unique(), lab_bounds$LOINC)

patient_obs_wide <- patient_obs %>% 
  left_join(lab_bounds, by = c('concept_code' = 'LOINC')) %>% 
  filter(days_since_admission >= 0, !is.na(short_name)) %>%
  select(- concept_code) %>% 
  pivot_wider(id_cols = c(patient_num, days_since_admission),
              names_from = short_name,
              values_from = value,
              values_fn = mean)  %>% 
  left_join(select(demo_raw, patient_num, severe), by = "patient_num") %>% 
  mutate(severity = factor(severe) %>% 
           fct_recode("severe" = "1", "nonsevere" = "0")) %>% 
  select(- severe)

lab_names <- intersect(lab_bounds$short_name, names(patient_obs_wide))
```


```{R}
te_labs <- patient_obs_wide %>% 
  filter(days_since_admission <= 20,
         days_since_admission >= 0) %>% 
  mutate(te = patient_num %in% unique(te_patients$patient_num)) %>% 
  select(- c(patient_num, days_since_admission))

vis_miss(te_labs, sort_miss = TRUE)

```

```{r}
gg_miss_upset(te_labs, nsets = 7, nintersects = 10)
gg_miss_var(te_labs) +
  theme(panel.grid.minor = element_blank(),
        panel.grid.major.y = element_blank()) + 
  scale_y_continuous(expand = expansion(0, c(0, 1000))) +
  xlab(NULL)

gg_miss_var(te_labs, show_pct = TRUE) +
  theme(panel.grid.minor = element_blank(),
        panel.grid.major.y = element_blank()) + 
  scale_y_continuous(expand = expansion(0, c(0, 1))) +
  xlab(NULL)
# mcar_test(te_labs)
```



```{R}
#check NAs in the Wide format
na_stats <- patient_obs_wide %>% 
  select(- c(patient_num, days_since_admission)) %>% 
  is.na() %>% 
  `!` 

na_df <- data.frame(value_existed = colSums(na_stats), 
           prop_existed = colMeans(na_stats)) %>% 
  rownames_to_column('lab') %>% 
  mutate(prop_na = 1 - prop_existed,
         lab = fct_reorder(lab, value_existed))

n_values <- na_df %>% 
  ggplot(aes(x = value_existed, y = lab)) +
  geom_col() + 
  labs(x = 'Number of values', y = NULL)

na_prob <- na_df %>%
  rename('Valid value' = prop_existed, 'NA' = prop_na) %>% 
  pivot_longer(c(`Valid value`, `NA`)) %>% 
  ggplot(aes(x = value, y = lab, fill = name)) +
  geom_col() + 
  scale_fill_discrete(guide = guide_legend(reverse = TRUE)) +
  labs(x = 'Proportion', y = NULL) +
  # guides(fill = guide_legend(reverse = TRUE))
  theme(
    axis.text.y = element_blank(),
    legend.key.width = unit(6, 'mm'),
    legend.key.height = unit(4, 'mm'),
    legend.position = 'bottom')

plot_grid(n_values, na_prob, nrow = 1, axis = 'b', align = 'h')

penn_na_df <- na_df
```


```{r}
patient_obs_long <- patient_obs_wide %>% 
  pivot_longer(- c(patient_num, days_since_admission, severity),
               names_to = 'lab', values_to = 'value',
               values_drop_na = TRUE)
```


```{r fig.show='hide'}
# number of patients per lab value

get_pat_labs <- function(labi){
  patients_labs <- patient_obs_long %>% 
  filter(lab == labi) %>% 
  count(patient_num) %>% 
  pull(n) %>% 
  hist(., breaks = seq(0, max(.), 1))

  patients_labs[[1]] <- patients_labs[[1]][-1] 
  data.frame(
    do.call(cbind.data.frame, patients_labs[1:3]),
    lab = labi)
}

patient_lab <- lapply(unique(patient_obs_long$lab), get_pat_labs) %>% 
  bind_rows()

penn_patient_lab <- patient_lab

melt_patient_lab <- patient_lab[rep(1:nrow(patient_lab), patient_lab$counts), ]
```

```{r}
melt_patient_lab %>% 
  filter(breaks < 30) %>% 
  mutate(lab = fct_infreq(lab)) %>% 
  ggplot(aes(x = breaks, y = lab, fill = lab, height = ..count..)) +
  # geom_density_ridges(stat = "identity", scale = 2) +
  geom_ridgeline(stat="binline", binwidth=1, scale = 0.001) +
  # scale_color_viridis_d(option = 'C') +
  scale_fill_viridis_d(option = 'C', guide = FALSE) +
  labs(y = NULL) +
  # coord_flip() +
  # facet_wrap(~ lab, scales = 'free') + 
  NULL
```


## Number of observation (days) per patient
```{R}
days_count_min_max <- 
  # multiple_imps %>%
  patient_obs_wide %>% 
  group_by(patient_num, severity) %>%
  summarise(
    n_values = n_distinct(days_since_admission),
    min_day = min(days_since_admission),
    max_day = max(days_since_admission),
    .groups = 'drop'
  ) %>% 
  # left_join(demo_raw, by = "patient_num") %>% 
  # mutate(time_obs = max_day - min_day,
  #        severity = factor(severe)) %>% 
  # select(-patient_num) %>% 
  add_count(severity, name = 'n_severity') %>%
  {.}

penn_agg_n_values <- days_count_min_max %>% 
  count(severity, n_severity, n_values,
        name = 'n_nvals')
penn_agg_max_day <- days_count_min_max %>% 
  count(severity, n_severity, max_day,
        name = 'n_maxday')

penn_agg_n_values %>% 
  ggplot(aes(x = n_nvals)) +
  geom_bar()

(n_severe <- sum(days_count_min_max$severity == 'severe'))
(n_nonsevere <- sum(days_count_min_max$severity == 'nonsevere'))
```

## Histogram of the number of days with at least one observation
```{R}
penn_agg_n_values %>% 
  ggplot(aes(x = n_values, y = n_nvals, fill = severity)) +
  geom_col(alpha = 0.5) +
  scale_fill_brewer(palette = 'Dark2', direction = -1) +
  labs(fill = 'Severe?', 
       x = "Number of days with data", 
       y = "Count")
```

## Histogram of length of stay
i.e. last day with observation-first day with observation

We need to check for readmission here.
```{R}
penn_agg_max_day %>% 
  ggplot(aes(x = max_day, y = n_maxday, fill = severity)) +
  geom_col(alpha = 0.5) +
  scale_fill_brewer(palette = 'Dark2', direction = -1) +
  labs(fill = 'Severe?', 
       x = "Number of days with data", 
       y = "Count")
```

## Analyze missingness and frequency of measures for each lab

```{r}
per_lab <- patient_obs_long %>% 
  group_by(lab, patient_num, severity) %>% 
  count(name = 'n_obs') %>% 
  ungroup() %>% 
  group_by(lab, severity) %>% 
  count(n_obs) %>% 
  ungroup() %>% 
  pivot_wider(names_from = severity, values_from = n, values_fill = 0) %>% 
  mutate(both_severities = nonsevere + severe) %>% 
  mutate(prop_nonsevere = nonsevere/n_nonsevere,
         prop_severe = severe/n_severe,
         prop_both = both_severities/nrow(days_count_min_max))

lab_medians <-
  patient_obs_long %>% 
  add_count(lab, name = 'total_obs') %>% 
  group_by(lab) %>% 
  mutate(total_patients = length(unique(patient_num))) %>% 
  add_count(patient_num, name = 'n_obs_patients') %>% 
  mutate(median_obs_per_patient = median(n_obs_patients),
         n_greater0 = n_obs_patients > median_obs_per_patient,
         n_greater1 = n_obs_patients > median_obs_per_patient + 1,
         n_greater2 = n_obs_patients > median_obs_per_patient + 2) %>% 
  group_by(severity) %>% 
  mutate(each_med_obs_per_patient = median(n_obs_patients)) %>% 
  ungroup(severity) %>% 
  select(- c(days_since_admission, value)) %>% 
  distinct() %>% 
  group_by(lab) %>% 
  mutate(across(contains('n_greater'), sum)) %>% 
  select(- c(n_obs_patients, patient_num)) %>% 
  distinct() %>% 
  pivot_wider(names_from = severity, values_from = each_med_obs_per_patient) %>% 
  rename('median_obs_per_severe_patient' = severe,
         'median_obs_per_non_severe_patient' = nonsevere) %>% 
  select(lab, total_obs, total_patients, starts_with('med'), starts_with('n_'))

lab_medians %>% 
  datatable(rownames = FALSE)
```


`n_greater0` shows the number of patients who had more observations than the median.
`n_greater1` shows the number of patients who had more observations than the median + 1.
`n_greater2` shows the number of patients who had more observations than the median + 2.

```{r eval=FALSE, include=FALSE}
lab_medians %>% 
  select(lab, total_obs, total_patients) %>%
  pivot_longer(- lab) %>% 
  ggplot(aes(x = value, y = fct_reorder(lab, value))) +
  geom_col() +
  facet_grid(cols = vars(name), scales = 'free_x', space = 'free') +
  labs(y = NULL)
```

In the figure below:

- Grey dash line: `Reference Low`
- Grey solid line: `Reference High`
- Black dash line: `lower bound outlier` (QC)
- Black solid line: `upper bound outlier` (QC)

```{r fig.width=9, fig.height=11}
patient_obs_long %>% 
  left_join(lab_bounds, by = c('lab' = 'short_name')) %>% 
  ggplot(aes(y = severity, x = value, fill = severity)) +
  geom_violin() +
  scale_fill_brewer(palette = 'Dark2', guide = guide_legend(reverse = TRUE)) +
  labs(y = NULL, x = NULL) +
  geom_vline(aes(xintercept = `Reference Low`), linetype = 'dashed', color = 'grey') +
  geom_vline(aes(xintercept = `Reference High`), color = 'grey') +
  geom_vline(aes(xintercept = LB), linetype = 'dashed') +
  geom_vline(aes(xintercept = UB)) +
  facet_wrap(~ lab, scales = 'free', ncol = 2, strip.position = 'left') +
  theme(axis.text.y = element_blank())
```

## Missing data heatmap

"Binned" heatmap

```{r fig.width=12}
per_lab %>%
  mutate(obs_bin = cut(
    n_obs, 
    breaks = c(0:15, 20, 30, max(n_obs)))) %>% 
  group_by(lab, obs_bin) %>% 
  summarise(both_severities = sum(both_severities), 
            severe = sum(severe), 
            nonsevere = sum(nonsevere), 
            .groups = 'drop') %>% 
  select(lab, obs_bin, both_severities, severe, nonsevere) %>% 
  pivot_longer(c(both_severities, severe, nonsevere)) %>% 
  mutate(name = name %>% fct_recode(
    'All patients' = 'both_severities',
    'Severe patients' = 'severe',
    'Non-severe patients' = 'nonsevere'
  )) %>% 
  ggplot(aes(x = obs_bin, fill = value, y = fct_reorder(lab, value))) + 
  geom_tile(colour = "white", size = 0.2) +
  geom_text(aes(label = value), colour = "white", size = 2) +
  scale_y_discrete(expand = c(0, 0))+
  scale_fill_gradient(low = "lightgrey", high = "darkblue") +
  facet_wrap(~ name, nrow = 1) +
  labs(x = 'Binned number of values a patient has for each lab',
       y = NULL, fill = '# patients') +
  theme(panel.grid.major = element_blank(),
        legend.position = c(0.93, 0.2),
        axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5),
        axis.ticks.y = element_blank()
  )
```


```{r fig.width=12}
per_lab %>%
  mutate(obs_bin = cut(n_obs, breaks = c(0:15, 20, 30, max(n_obs)))) %>% 
  group_by(lab, obs_bin) %>% 
  summarise(prop_both = sum(prop_both, na.rm = TRUE), 
            prop_severe = sum(prop_severe, na.rm = TRUE), 
            prop_nonsevere = sum(prop_nonsevere, na.rm = TRUE), 
            .groups = 'drop') %>% 
  select(lab, obs_bin, prop_both, prop_severe, prop_nonsevere) %>% 
  pivot_longer(c(prop_both, prop_severe, prop_nonsevere)) %>% 
  mutate(name = name %>% fct_recode(
    'Compared to all patients' = 'prop_both',
    'Compared to all severe patients' = 'prop_severe',
    'Compared to all non-severe patients' = 'prop_nonsevere'
  )) %>% 
  ggplot(aes(x = obs_bin, fill = value, y = fct_reorder(lab, value))) + 
  geom_tile(colour = "white", size = 0.2) +
  geom_text(aes(label = round(value, 2)), colour = "white", size = 2) +
  scale_y_discrete(expand = c(0, 0)) +
  scale_fill_gradient(low = "lightgrey", high = "darkblue",
                      labels = scales::percent_format(accuracy = 1L)) +
  facet_wrap(~ name, nrow = 1) +
  labs(x = 'Binned number of values a patient has for each lab',
       y = NULL, fill = '% patients') +
  theme(panel.grid.major = element_blank(),
        legend.position = c(0.93, 0.2),
        axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5),
        axis.ticks.y = element_blank()
  )
```

Denominator: total number of patients, total number of non-severe patients,
and total number of severe patients, respectively.

```{r}
per_lab %>% 
  select(lab, n_obs, severe, nonsevere) %>% 
  filter(n_obs <= 90) %>%
  pivot_longer(c(severe, nonsevere)) %>% 
  mutate(lab = fct_reorder(lab, n_obs),
         name = name %>% fct_recode(
    'Severe patients' = 'severe',
    'Non-severe patients' = 'nonsevere'
  )) %>% 
  ggplot(aes(x = n_obs, fill = name, y = value)) + 
  geom_col() +
  scale_fill_brewer(palette = 'Dark2', direction = -1) +
  facet_wrap(~ lab, scales = 'free') +
  labs(x = 'Number of values a patient has for each lab',
       y = NULL, fill = '# patients') +
  theme(panel.grid.major = element_blank(),
        legend.position = c(0.9, 0.2),
        axis.ticks.y = element_blank()
  )
```


```{r}
sessionInfo()
```

```{r}
save(penn_na_df, penn_agg_n_values, penn_agg_max_day, penn_patient_lab,
     # multiple_imps,
     file = 'results/penn-results.Rdata')
```

